---
title: "Project - Wenxuan Sun"
author: "Wenxuan Sun"
date: "09/04/2021"
output:
  html_document: default
  pdf_document: default
---

# Dataset information:
- data_full.RData contains data of 928 stocks from 1995-01-31 to 2021-02-26, updated every month.

# Steps:
- Set up a benchmark (data_index.RData) by attributing 100 USD to each stock (928 in total) at the beginning date 1995-01-31. The amount bought at that moment is calculated by 100USD/Close price at the beginning.
- Set up a trading strategy: select the stocks with Mkt_Cap quantile in 40%~85%. Then attribute equal amount of USD to each of them.
- Compare the performance of index and the portfolio


Load required packages
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(tibble)
library(tseries)
```

Set workpath and load datasets
```{r}
setwd("C:/Gggggg/MSc in Finance EMLYON/Data Science R/Project")
load("data_esg.RData")
load("data_full.RData")
head(data_full)
summary(data_full)
```
Function to change the positions
```{r}
# Warning: this only applies to equal USD amount attributed to each stock
# The target stock will be bought in equal USD amount with available cash
change_pos = function(date, pos_now, tick_tgt){
  # date of changing position should be in date format "yyyy-mm-dd"
  # pos_now should be a 2column*nrow dataframe containing $Tick and $Position
  # tick_tgt should be a vector or a list of Tick (only the ones available in data_full)
  df_now = data_full[data_full$Date == date, ]
  price_now = df_now %>% filter(Tick %in% pos_now$Tick) %>% select(Close)
  cash = t(price_now) %*% pos_now1$Position
  price_tgt = df_now %>% filter(Tick %in% tick_tgt) %>% select(Close)
  pos_tgt = rep(cash / length(tick_tgt), length(tick_tgt)) / price_tgt$Close
  return(data.frame(Tick=tick_tgt, Position=pos_tgt))
}
```
Set up the index table (benchmark)
The index is built with 100 USD for each stock at the beginning date 1995-01-31.
The index at the first date is 92,800 USD and ends at 2,475,168.2 USD on 2021-02-26.
```{r}
data_full = data_full %>% arrange(Date)  # arrange dates
data_full$Date = as.Date(data_full$Date, origin="1970-01-01")

min_date = min(data_full$Date)  # begin date
max_date = max(data_full$Date)  # end date

c100 = rep(100, length(data_full[data_full$Date == min_date, ]$Close))  # create a vector with length of 928 filling with 100

close_1 = data_full[data_full$Date == min_date, ]$Close # close price of 928 stocks at the beginning date
Position = c100 / close_1  # positions for 928 stocks

# index = position * price
index_1 = t(Position) %*% close_1  # index at the beginning

data_index = data.frame(Date=character(), Index=numeric())  # dataframe of the index

for(i in unique(data_full$Date)){
  i = as.Date(i, origin="1970-01-01")
  close_i = data_full[data_full$Date == i, ]$Close
  index_i = t(Position) %*% close_i
  data_index_i = data.frame(Date=i, Index=index_i)
  data_index = rbind(data_index, data_index_i)
}


# Add a return column in data_index
data_index = data_index %>%
  mutate(Index_Return = Index / lag(Index) - 1)
head(data_index)

# remove variables that are useless later
rm(c100)
rm(close_1)
rm(close_i)
rm(i)
rm(Position)
rm(index_i)
rm(index_1)
rm(data_index_i)
```
Set up a portfolio at the percentile 40%~85% of Market Cap
If the value of portfolio goes below 90% of index value, the percentile of Mkt_Cap will be recalculated and a new portfolio will be selected.
```{r}
# data at the beginning date 1995-01-31
data_full_1 = data_full %>% 
  filter(Date == min_date) %>% 
  arrange(Tick)
data_full_1 = data_full_1[!is.na(data_full_1$Mkt_Cap), ]  # remove na values in market cap column

quan_upper_1 = unname(quantile(data_full_1$Mkt_Cap, probs=0.85, na.rm=TRUE))  # 85% quantile of market cap on 1995-01-31
quan_lower_1 = unname(quantile(data_full_1$Mkt_Cap, probs=0.40, na.rm=TRUE))  # 40% quantile of market cap on 1995-01-31
close_1 = data_full_1[data_full_1$Mkt_Cap <= quan_upper_1 & data_full_1$Mkt_Cap >= quan_lower_1, ]$Close  # Close price of the selected stocks on 1995-01-31
stock_selected = data_full_1[data_full_1$Mkt_Cap <= quan_upper_1 & data_full_1$Mkt_Cap >= quan_lower_1, ]$Tick  # list of selected stock
money = rep(length(unique(data_full$Tick)) * 100 / length(stock_selected), length(stock_selected))  # money attributed to each selected stock
position = money / close_1  # Positions of selected stock

data_port = data.frame(Date=character(), Port_Value=numeric)  # dataframe of the portfolio
data_selected = data_full %>% filter(Tick %in% stock_selected)  # data of selected stocks

print(length(stock_selected))

for(i in unique(data_selected$Date)){
  i = as.Date(i, origin="1970-01-01")
  close_i = data_selected[data_selected$Date == i, ]$Close
  value_i = t(position) %*% close_i
  value_index_i = data_index[data_index$Date == i, ]$Index
  data_port_i = data.frame(Date=i, Port_Value=value_i)
  data_port = rbind(data_port, data_port_i)
}

# Add a return column in data_port
data_port = data_port %>%
  mutate(Port_Return = Port_Value / lag(Port_Value) - 1)
head(data_port)

# Remove unnecessary variables
rm(data_full_1)
rm(data_port_i)
rm(data_selected)
rm(value_i)
rm(close_1)
rm(close_i)
rm(i)
rm(money)
rm(position)
rm(quan_upper_1)
rm(quan_lower_1)
rm(stock_selected)

# for(i in unique(data_full$Date)){
#   i = as.Date(i, origin="1970-01-01")  # convert i into date format
#   data_full_i = data_full %>% filter(Date == i)
#   quan_90_i = unname(quantile(data_full_i$Mkt_Cap, probs=0.9, na.rm=TRUE))
#   portfolio_i = data_full_i %>% filter(Mkt_Cap >= quan_90_i)
# }
```
Plot the trend of benchmark and the portfolio
```{r}
data_trend = merge(data_port, data_index, by="Date") %>%
  subset(select=-c(Port_Return, Index_Return)) %>% 
  reshape2::melt(id.var='Date')
data_trend %>% 
  ggplot(aes(x=Date, y=value, color=variable)) +  
  geom_line() +
  ggtitle("Portfolio vs. Index") +
  xlab("Date") + 
  ylab("Value in USD") +
  theme(plot.title = element_text(color="red", size=18, face="bold.italic"),
    axis.title.x = element_text(color="green", size=16, face="bold"),
    axis.title.y = element_text(color="#FF11BB", size=16, face="bold"))
rm(data_trend)

data_return = merge(data_port, data_index, by="Date") %>% 
  subset(select=-c(Port_Value, Index)) %>% 
  reshape2::melt(id.var='Date') %>% 
  na.omit()
# Simple monthly return
data_return %>% 
  ggplot(aes(x=Date, y=value)) +  
  geom_area(aes(color=variable), alpha=0.4) +
  ggtitle("Portfolio return vs. Index return") +
  xlab("Date") + 
  ylab("Return") +
  labs(caption="This is the simple monthly return calculated by division") +
  theme(plot.title = element_text(color="red", size=18, face="bold.italic"),
    axis.title.x = element_text(color="green", size=16, face="bold"),
    axis.title.y = element_text(color="#FF11BB", size=16, face="bold"),
    plot.caption = element_text(color="#22AAFF", size=12))
```
Compare the performance between portfolio and index
```{r}
# total return
total_rt_index = data_index[data_index$Date == max_date, ]$Index / 
  data_index[data_index$Date == min_date, ]$Index - 1  # total return of index
total_rt_port = data_port[data_port$Date == max_date, ]$Port_Value / 
  data_port[data_port$Date == min_date, ]$Port_Value - 1  # total return of index

# Index return volatility (SD)
rt_vol_index = 12 * sd(data_index$Index_Return, na.rm=TRUE)
# Portfolio return volatility (SD)
rt_vol_port = 12 * sd(data_port$Port_Return, na.rm=TRUE)

# Maximum drawdown
max_dd_index = tseries::maxdrawdown(data_index$Index)$maxdrawdown
max_dd_port = tseries::maxdrawdown(data_port$Port_Value)$maxdrawdown

summary_df = data.frame(Type=c("Index", "Portfolio"),
                        Value_Begin=c(data_index[data_index$Date == min_date, ]$Index, 
                                      data_port[data_port$Date == min_date, ]$Port_Value),
                        Value_End=c(data_index[data_index$Date == max_date, ]$Index,
                                    data_port[data_port$Date == max_date, ]$Port_Value),
                        Total_Return=c(total_rt_index, total_rt_port),
                        Return_Volatility=c(rt_vol_index, rt_vol_port),
                        Maximum_Drawdown=c(max_dd_index, max_dd_port))
print(summary_df)
```



